\Large \bf A first look at auditing in a Blockchain world

### 2018-09-24 auditing/Latex/sections/01-intro.tex
[auditing] edit and citations till 2.3

When assessing whether or not to take on a new client, auditors lack experience in this sector and therefore are unable to develop expectations of financial performance as a way of challenging financial statement assumptions. Due to the complex and rapidly changing technological environment, auditors are also unable to keep pace with the changes and develop the in-depth knowledge of their clients\textquotesingle  businesses required for performing an audit. Finally, auditors are wary of accepting clients that hold a significant amount of cryptoassets as this space is largely unregulated. A lack of third-party oversight puts an onus on the auditor, further increasing their risk exposure.

@@

\paragraph{\textbf{Methodology.}} To ensure a comprehensive overview of the changes facing auditors of companies deploying blockchain technologies, we first used structured brainstorming within our multidisciplinary research team, which includes expertise in both auditing and blockchain technologies. Once our preliminary list of challenges were established, we vetted our results through informal discussions with over a dozen individuals working in the field, including at the Big Four and mid-sized accounting firms, and used these interviews to augment our list. The contribution of this paper is to provide a comprehensive list of challenges, rather than determining the relative importance of each or any broader concepts across the industry---thus we did not find it necessary to apply qualitative data analysis (\eg grounded theory) to the interviews; instead, we simply extracted the challenges raised which we initially missed.

\paragraph{\textbf{Relevance.}} Our work can viewed as a case study of cryptographers working with an outside profession for addressing a real-world problem. The result is not a new protocol but a two-way knowledge transfer between professions that we hope is interesting to this audience. While displacing auditors is occasionally a target of cryptographers~\cite{narula2018zkledger,dagher2015provisions}, auditing itself is also occasionally studied directly at venues like Financial Cryptography

@@

\section{Preliminaries and Related Work}
Financial statements are prepared by a firm to summarize its assets and liabilities at year-end, the resulting changes in firm equity, and the firm’s income/profits, expenses/losses, and cash flows across the year. Because the firm generally wants to present the most optimistic view of its financial health, the firm may hire an auditor to ensure it is a realistic picture of the company's financial performance. In doing so, auditors are expected to comply with Generally Accepted Auditing Standards (GAAS) and the Code of Conduct of their professional order. In addition to an audited statement, the firm will produce quarterly statements that are reviewed by a third party auditor to a less rigorous standard.

\todo{Note that there are auditing and accounting papers in the bib/zotero that have not been cited here!}

%------------------------------------------------------------------------------------------------
\subsection{Primary Stakeholders in an Audit}

\paragraph{Firm Management.} Financial statements are prepared by and remain the responsibility of firm management. Many assumptions and estimates are required to prepare this report. A firm's CEO and CFO must certify the adequacy and effectiveness of the firm's internal controls over financial reporting, which poses a reputational risk for firm directors as they will be held accountable for financial statements that are misstated. Management will seek an audit to comply with regulatory requirements, or because they feel it will give them a competitive advantage over non-audited firms in raising capital and lowering their lending rates. In a market where auditors are reluctant or audits are expensive, firms can use their ability to obtain audited financial statements as a barrier against new entrants to the blockchain sector.

@@

\paragraph{Investors.} The primary users of financial statements are external investors who make decisions about whether or not to lend money or invest capital in a target firm. Investors use financial statements to assess their potential return on investment. Although \textit{caveat emptor} is an important consideration, investors rely on the existence of an auditor's report as a signal of the reliability of financial disclosures.

\paragraph{Financial Regulators.} Regulators span different government agencies with a spectrum of concerns including financial fraud, taxation, anti-money laundering, and know your customer rules. Financial audits are central to the role of security regulators, who require publicly traded firms to obtain an annual unqualified audit opinion on their financial statements for consumer protection. Failure to do so, would put the company offside with the regulators\textquotesingle  requirements and could result in the company\textquotesingle s shares being placed on cease-trade.

%------------------------------------------------------------------------------------------------
\subsection{Cryptoassets and firms that hold them}
The central component of a financial report is the firm’s balance sheet which lists assets and liabilities. We use the term \textit{cryptoasset} and \textit{cryptoliability} to refer to items whose value are contingent on blockchain technologies. Cryptoassets include cryptocurrencies and other tokens of tradeable value. Without loss of generality, we assume cryptocurrencies are native to their underlying blockchain, can be transacted directly, and are used to pay fees for transaction execution. Bitcoin and Ether are examples. For blockchains that allow developers to deploy custom decentralized apps, the apps might issue and manage the ownership of a custom tokens. In the case of Ethereum, these tokens are often ERC20 tokens, where ERC20 is sometimes misunderstood to mean what the token represents; rather, ERC20 is a technical standard about how token transactions are invoked.

What a token represents varies across applications but includes at least of the following major categories. (1) Access tokens: a service is developed which requires its own custom tokens for using the service; (2) Backed tokens: a token issuer claims to be holding something valuable (material or digital) in reserve, and the token represents a claim on these reserves; (3) Equity tokens: a firm issues tokens to represent ownership shares of the company; and (4) Collectable tokens: the token itself is offered as a contemporary collector’s item. Digital tokens of these types predate blockchains; for example, (1) Linden dollars, (2) E-Gold, and (4) URLs are arguably examples. Equities (3) are almost entirely dematerialized (\eg paper stock certificates are rare) but operate closer to (2) with a central depository.

Tokens and cryptocurrencies are issued to an initial set of owners through any mechanism of the issuer’s choosing. A popular option is auctioning a set of tokens to the public in an ``initial coin offering'' or ICO. This is intended to resemble the initial public offering (IPO) of a firm’s stock, but ICOs often lack any consumer protection or regulatory compliance. An ICO of an access token might raise capital for developing or improving the service that will use these tokens. Buyers obtain such tokens to lock in the purchase price (hedging) or for speculation.

Any action that results in a firm borrowing a cryptoasset results in an offsetting cryptoliability. For example, an exchange service that holds bitcoin or tokens on behalf of its clients has both an asset (the cryptoasset it holds in its possession) and a liability (the obligation to repay its clients). Tokens could represent a debt instrument, like a bond, commercial paper, or repurchase agreement. In a recent pilot, a blockchain-based certificate of deposit was issued. This creates an asset for the investors (Western Assets, Pfitzer, etc.) and a liability for the issuer (National Bank of Canada with J.P. Morgan).

@@

To illustrate how a firm might come to possess a cryptoasset, consider four examples of firms that have actually sought audits or regulatory exemptions from audits\footnote{[Anonymized] Based on data provided to us by the financial regulator in the region where the authors work.}. A mining firm will invest in specialized computing equipment and electricity to generate cryptocurrencies it will hold as assets. An exchange service will hold demand deposits of cryptocurrencies or tokens, as well as governmental currencies, and allow its users to trade them. An investment fund will hold a portfolio of cryptoassets (\eg TIQ101-CF is 50\% bitcoin, 35\% ether, 15\% litecoin \todo{[c]: look at the comment below in latex}) and sell shares to investors through a standard financial platform. Finally a firm initiating a token sale will raise capital in cryptocurrencies, held as an asset, in return for tokens that are not generally liabilities (it may also reserve some tokens as an additional asset).

@@

%------------------------------------------------------------------------------------------------
\subsection{Presentation of Cryptoassets}
A financial report will classify assets and liabilities according to categories established by the auditing standards adhered to in the firm\textquotesingle s reporting country (American Generally Accepted Accounting Principles in the United States and International Financial Reporting Standards (IFRS) in many other countries). The following discussion will apply IFRS although the considerations under US GAAP are similar. \todo{ GAAP and IFRS citations  }

@@

The Bitcoin, on the other hand, would be presented as an intangible asset. This investment could either be presented at historical cost (what the company paid for the Bitcoin) or revalued to its fair market value. If the Bitcoin is reported at cost, any increases in value would not be recorded until the asset is sold (decreases in value below cost would be reported immediately). If the Bitcoin is reported at fair market value, any increases in value would be reported in Other Comprehensive Income, a special category that does not fall into net profit. This means that while the investment fund holds two investments, Bitcoin and Google stock, both for long-term growth purposes, the current standards would not allow these investments to recorded on the same basis. This also means that increases in value of Bitcoin or other cryptoassets are not recorded as part of net profit, which obfuscates a company\textquotesingle s annual reported earnings and undermines their usefulness for investors.

>>>>

When assessing whether or not to take on a new client, auditors lack experience in this sector and therefore are unable to develop expectations of financial performance as a way of challenging financial statement assumptions. Due to the complex and rapidly changing technological environment, auditors are also unable to keep pace with the changes and develop the in-depth knowledge of their clients{'}  businesses required for performing an audit. Finally, auditors are wary of accepting clients that hold a significant amount of cryptoassets as this space is largely unregulated. A lack of third-party oversight puts an onus on the auditor, further increasing their risk exposure.

@@

\paragraph{\textbf{Methodology.}} To ensure a comprehensive overview of the changes facing auditors of companies deploying blockchain technologies, we first used structured brainstorming within our multidisciplinary research team, which includes expertise in both auditing and blockchain technologies. Once our preliminary list of challenges was established, we vetted our results through informal discussions with over a dozen individuals working in the field, including at the Big Four\footnote{The Big Four are the four biggest auditing firms in the world which includes, Ernst \& Young (EY), Deloitte \& Touche, KPMG and PricewaterhouseCoopers (PwC).} and mid-sized accounting firms, and used these interviews to augment our list. The contribution of this paper is to provide a comprehensive list of challenges, rather than determining the relative importance of each or any broader concepts across the industry---thus we did not find it necessary to apply qualitative data analysis (\eg grounded theory) to the interviews; instead, we simply extracted the challenges raised which we initially missed.

\paragraph{\textbf{Relevance.}} Our work can be viewed as a case study of cryptographers working with an outside profession for addressing a real-world problem. The result is not a new protocol but a two-way knowledge transfer between professions. While displacing auditors is occasionally a target of cryptographers~\cite{narula2018zkledger,dagher2015provisions}, auditing itself is also occasionally studied directly at venues like Financial Cryptography~\cite{grigg2000financial}.

@@

\section{Preliminaries and Related Work}
Financial statements are prepared by a firm to summarize its assets and liabilities at year-end, the resulting changes in firm equity, and the firm{'}s revenue, expenses, and cash flows across the year. Because the firm generally wants to present the most optimistic view of its financial health, the firm may hire an auditor to ensure it is a realistic picture of the company{'}s financial performance. In doing so, auditors are expected to comply with Generally Accepted Auditing Standards (GAAS) and the Code of Conduct of their professional order~\cite{strother1975establishment}. In addition to an audited statement, the firm will produce quarterly statements that are reviewed by a third party auditor to a less rigorous standard.

%NOTE: Note that there are auditing and accounting papers in the bib/zotero that have not been cited here! {blaze1998compliance}
%------------------------------------------------------------------------------------------------
\subsection{Primary Stakeholders in an Audit}

\paragraph{Firm Management.} Financial statements are prepared by and remain the responsibility of firm management. Many assumptions and estimates are required to prepare this report. A firm's CEO and CFO must certify the adequacy and effectiveness of the firm{'}s internal controls over financial reporting, which poses a reputational risk for firm directors as they will be held accountable for financial statements that are misstated. Management will seek an audit to comply with regulatory requirements, or because they feel it will give them a competitive advantage over non-audited firms in raising capital and lowering their lending rates. In a market where auditors are reluctant or audits are expensive, firms can use their ability to obtain audited financial statements as a barrier against new entrants to the blockchain sector.

@@

\paragraph{Investors.} The primary users of financial statements are external investors who make decisions about whether or not to lend money or invest capital in a target firm. Investors use financial statements to assess their potential return on investment. Although \textit{caveat emptor} is an important consideration, investors rely on the existence of an auditor{'}s report as a signal of the reliability of financial disclosures.

\paragraph{Financial Regulators.} Regulators span different government agencies with a spectrum of concerns including financial fraud, taxation, anti-money laundering, and know your customer (KYC) rules. Financial audits are central to the role of security regulators, who require publicly traded firms to obtain an annual unqualified audit opinion on their financial statements for consumer protection. Failure to do so, would put the company offside with the regulators{'} requirements and could result in the company{'}s shares being placed on cease-trade.

%------------------------------------------------------------------------------------------------
\subsection{Cryptoassets and firms that hold them}
The central component of a financial report is the firm{’}s balance sheet which lists assets and liabilities. We use the term \textit{cryptoasset} and \textit{cryptoliability} to refer to items whose value are contingent on blockchain technologies. Cryptoassets include cryptocurrencies and other tokens of tradeable value. Without loss of generality, we assume cryptocurrencies are native to their underlying blockchain, can be transacted directly, and are used to pay fees for transaction execution, \eg Bitcoin and Ether. For blockchains that allow developers to deploy custom decentralized applications, the applications might issue and manage the ownership of custom tokens. In the case of Ethereum, these tokens are often \textit{ERC20} tokens~\cite{erc20}, where \textit{ERC20} is sometimes misunderstood to mean what the token represents; rather, \textit{ERC20} is a technical standard about how token transactions are invoked.

What a token represents varies across applications but includes at least one of the following major categories. \textit{(1) Access tokens}: a service is developed which requires its own custom tokens for using the service; \textit{(2) Backed tokens}: a token issuer claims to be holding something valuable (material or digital) in reserve, and the token represents a claim on these reserves; \textit{(3) Equity tokens}: a firm issues tokens to represent ownership shares of the company; and \textit{(4) Collectable tokens}: the token itself is offered as a contemporary collector{’}s item (\eg \textit{ERC721}~\cite{erc721}). Digital tokens of these types predate blockchains; for example, \textit{Linden dollars}, \textit{E-Gold}, and \textit{(4)} URLs are arguably examples. Equities \textit{(3)} are almost entirely dematerialized (\eg paper stock certificates are rare) but operate closer to \textit{(2)} with a central depository.

Tokens and cryptocurrencies are issued to an initial set of owners through any mechanism of the issuer{'}s choosing. A popular option is auctioning a set of tokens to the public in an ``Initial Coin Offering'' or ICO. This is intended to resemble the ``Initial Public Offering'' (IPO) of a firm{'}s stock, but ICOs often lack any consumer protection or regulatory compliance. An ICO of an access token might raise capital for developing or improving the service that will use these tokens. Buyers obtain such tokens to lock in the purchase price (hedging) or for speculation.

Any action that results in a firm borrowing a cryptoasset results in an offsetting cryptoliability. For example, an exchange service that holds bitcoin or tokens on behalf of its clients has both an asset (the cryptoasset it holds in its possession) and a liability (the obligation to repay its clients). Tokens could represent a debt instrument, like a bond, commercial paper, or repurchase agreement. In a recent pilot, a blockchain-based certificate of deposit was issued~\cite{NBCJPMorganBlockchain}. This creates an asset for the investors (Western Assets, Pfitzer, etc.) and a liability for the issuer (National Bank of Canada with J.P. Morgan).

@@

To illustrate how a firm might come to possess a cryptoasset, consider four examples of firms that have actually sought audits or regulatory exemptions from audits\footnote{[Anonymized] Based on data provided to us by the financial regulator in the region where the authors work.}. A mining firm will invest in specialized computing equipment and electricity to generate cryptocurrencies it will hold as assets. An exchange service will hold demand deposits of cryptocurrencies or tokens, as well as governmental currencies, and allow its users to trade them. An investment fund will hold a portfolio of cryptoassets (\eg TIQ101-CF is 50\% bitcoin, 35\% ether, 15\% litecoin \todo{[cite-confidential\?]}) and sell shares to investors through a standard financial platform. Finally, a firm initiating a token sale will raise capital in cryptocurrencies, held as an asset, in return for tokens that are not generally liabilities (it may also reserve some tokens as an additional asset).

@@

%------------------------------------------------------------------------------------------------
\subsection{Presentation of Cryptoassets}
A financial report will classify assets and liabilities according to categories established by the auditing standards adhered to in the firm{'}s reporting country (American Generally Accepted Accounting Principles in the United States and International Financial Reporting Standards (IFRS) in many other countries). The following discussion will apply IFRS although the considerations under US GAAP are similar. \todo{ GAAP and IFRS citations  }

@@

The Bitcoin, on the other hand, would be presented as an intangible asset. This investment could either be presented at historical cost (what the company paid for the Bitcoin) or revalued to its fair market value. If the Bitcoin is reported at cost, any increases in value would not be recorded until the asset is sold (decreases in value below cost would be reported immediately). If the Bitcoin is reported at fair market value, any increases in value would be reported in Other Comprehensive Income, a special category that does not fall into net profit. This means that while the investment fund holds two investments, Bitcoin and Google stock, both for long-term growth purposes, the current standards would not allow these investments to recorded on the same basis. This also means that increases in value of Bitcoin or other cryptoassets are not recorded as part of net profit, which obfuscates a company{'}s annual reported earnings and undermines their usefulness for investors.
